---
author: |
  - |
    Mbanguta Lwando
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
classoption: notitlepage
geometry: margin=1in
header-includes:
  - "\\usepackage{graphicx}"
  - \usepackage{pdfpages}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Topic 2: Introduction to Alternative Investments

## SubTopic 2.2: Environment of Alternative Investments

### Participants

Can be divided into four major categories:

#### 1) Buy Side

##### 

Refers to the institutions and entities that buy large quantities of securities for the portfolios they manage.

It contrasts with the sell side, which focuses on distributing securities to the public.

##### Institutional Allocators

- Plan sponsors

A designated party, such as a company or an employer, that establishes a health care or retirement plan (pension) for employees that has special legal or taxation status.

Are companies or other collectives that establish the health care and retirement plans for the benefit of the organization's employees or members.

These funds can be designed as defined contribution plans or defined benefit plans. 

In defined contribution plans, employees take the investment risk, while that risk is borne by the employer in defined benefit plans. 

Defined benefit plans are more commonly associated with investing in and allocating to alternative investments.

- Foundations and endowments

A foundation is a not‐for‐profit organization that donates funds and support to other organizations for its own charitable purposes. 

An endowment is a fund bestowed on an individual or institution to be used by that entity for specific purposes and with principal preservation in mind.

- Family office and private wealth

Are private management advisory firms that serve ultra‐high‐net‐worth investors. 

A family office is a group of investors joined by familial or other ties who manage their personal investments as a single entity, hiring professionals to manage money for members of the office.

- Sovereign wealth funds

Are state‐owned investment funds held for the purpose of future generations and/or to stabilize the state currency.

These funds may emanate from budgetary and trade surpluses, perhaps through exportation of natural resources and raw materials.

Are accumulated to help provide financial stability and future opportunities for citizens and governments.

##### Investment Management Structures

- Private limited partnerships

Are a form of business organization that potentially offers the benefit of limited liability to the organization's limited partners but not to its general partner.

For tax purposes, limited partnerships tend to flow taxable revenue and expenses directly through to their partners rather than being taxed at the partnership level.

- Private investment pools

Focus on serving as intermediaries between investors and alternative investments.

Limited partnerships are usually managed by the general partner, while most of the invested funds are provided by the limited partners.

- Separately managed accounts (SMAs)

Are individual investment accounts offered by a brokerage firm and managed by independent investment management firms.

The relationship between an investment adviser and a client to whom advice is provided is typically documented by a written investment management agreement.

A customer pays a fee to a money manager for managing the customer's investment (similar to pooled investment arrangements), but different in the following ways:

i) A fund investor owns shares of a company that in turn owns other investments, whereas an SMA investor actually owns the invested assets as the owner on record.

ii) A fund invests for the common purposes of multiple investors, whereas an SMA may have objectives tailored to suit the specific needs of its only investor.

iii) A fund is often opaque to its investors to promote confidentiality; an SMA offers transparency to its investor.

iv) Whereas the fund structure may allow investors to have limited liability, the SMA format may allow losses to be greater than the capital contribution when leverage or derivatives are used.

- Mutual funds (’40 Act funds)

Are registered investment pools offering their shareholders pro rata claims on the fund's portfolio of assets.

- UCITS (Undertakings for Collective Investment in Transferable Securities)

Allow retail access and marketing of hedge‐fund‐like investment pools.

- Master limited partnerships (MLPs)

Are publicly traded investment pools that are structured as limited partnerships and that offer their owners pro rata claims.

MLP units are traded on major stock exchanges, but have legal and tax structures similar to those of private limited partnerships.

#### 2) Sell Side

#####

Make their research available to their clients and are more focused on facilitating transactions than on managing money.

- Large dealer banks

Are major financial institutions that deal in securities and derivatives.

Act as intermediaries in the markets for securities, repurchase agreements, securities lending, and over‐the‐counter (OTC) derivatives.

Are often engaged in proprietary trading.

Have large asset management divisions that cater to the investment management needs of institutional and wealthy individual clients.

Operate internal hedge funds and take on private equity partnerships as part of their business management service.

The bank acts as a general partner with limited‐partner clients.

The role of dealer banks in the primary market is to intermediate between issuers and investors, to provide liquidity, and to act as underwriters of investments. 

In secondary securities markets, large dealer banks trade with one another and with brokers/dealers directly over the computer or the phone, as well as play an intermediating role of facilitating trades.

Also engage in proprietary trading.

Proprietary trading occurs when a firm trades securities with its own money in order to make a profit.

Serve as counterparties to OTC derivatives such as options, forwards, and swaps that require the participation of a counterparty dealer who meets customer demand by taking the opposite side of a desired position.

Are active as prime brokers that offer professional services specifically to hedge funds and other large institutional customers.

Have ventured into off‐balance‐sheet financing methods, a practice that involves a form of accounting in which large expenditures are kept off the company's balance sheet through various classification methods. 

Companies use off‐balance‐sheet financing to keep their debt‐to‐equity and leverage ratios low.

They share many of the same responsibilities as conventional commercial banks, including deposit taking and lending to corporations and consumers.

- Brokers

Receive commissions for executing transactions and have research departments that make investment recommendations.

Advantages of using brokers include their expertise in the trading process, their access to other traders and exchanges, and their ability to facilitate clearance and settlement.

Traders can use broker services when they want to remain anonymous to other traders.

Traders can manage their order exposure by breaking up large trades and distributing them to different brokers or by asking a single broker in charge of the entire trade to expose only parts of the order, so that the full size remains unknown to other traders.

The brokerage firm's proprietary trading operations involve the firm's own account, called the house account. 

Other sources of broker revenue include soft commissions, payments for order flow, interest on margin loans, short interest rebates (on short sales), underwriting fees when the firm helps clients sell securities, and mergers and acquisitions (M&A) fees. 

The major cost of running a brokerage firm is labour.

Brokerage firms and other firms with major investment activities organize their activities into three major operations: 

i) Front office operations involve investment decision‐making and contact with clients.

ii) Back office operations play a supportive role in the maintenance of accounts and information systems used to transmit important market and trader information in all trading transactions, as well as in the clearance and settlement of the trades.

iii) Middle office operations form the interface between the front office and the back office, with a focus on risk management.

#### 3) Outside Service Providers

#####

Alternative investment funds rely on outside service providers for their successful creation and operation.

- Prime brokers

Allow an investment manager to carry out trades in multiple financial instruments at multiple broker‐dealers while keeping all cash and securities at a single firm.

Have primary functions: clearing and financing trades for clients, providing research, arranging financing, borrowing and lending securities, and producing portfolio accounting

Have a powerful tool in their ability to make margin calls.

Can demand that the fund manager deposit more cash into its prime brokerage account to support its leveraged trading, and that the fund manager liquidate outstanding portfolio positions to raise cash to deposit with the prime broker.

The ultimate threat is that the prime broker can seize collateral from the hedge fund manager and liquidate the collateral to raise cash.

- Accountants and auditors

The accounting firm providing services to a hedge fund/another alternative investment fund should include an experienced auditor and tax adviser. 

During the creation of the fund or investment vehicle, the accounting firm provides services largely parallel to those of an attorney: reviewing legal documents to ensure that accounting methods and allocations are appropriate and feasible, and that relevant tax issues have been addressed. 

The accountant helps prepare partnership returns and the necessary forms for the investors in the fund to report their shares of partnership income, deductions, gains, and losses.

The adviser also provides tax‐related advice to the fund throughout the year and may be called on as a consultant on structuring and compensation issues for the principals of the general partner. 

The auditor performs a year‐end audit of the fund, including the review of security pricing, and presents the results of this audit to the fund and its investors.

- Attorneys

Helps determine the best legal structure for a fund's unique investment strategies, objectives, and desired investors. 

Takes care of filing any documents required by the government and creates the legal documents necessary for establishing and managing a hedge fund or another alternative investment. 

Can offer guidance on marketing a hedge fund or another alternative investment in full compliance with all legal requirements, as well as on operational issues, such as personal trading.

- Fund administrators

Responsible for bookkeeping, third‐party information gathering, and securities valuation functions for all of their funds, both onshore and offshore.

Has ten roles:

i) maintains a general ledger account

ii) marks the fund's books

iii) maintains its records

iv) carries out monthly accounting

v) supplies its monthly P&L statements and calculates its returns

vi) verifies asset existence

vii) independently calculates fees

viii) provides an unbiased, third‐party resource for price confirmation on security positions

ix) produces a monthly capital account statement for investors, and

x) apportions fund income or loss among them

Takes over the duties of day‐to‐day accounting and bookkeeping so that managers can focus on maximizing the portfolio's returns.

- Hedge fund infrastructure

May have three main financial components:

i) Financial platforms are systems that provide access to financial markets, portfolio management systems, accounting and reporting systems, and risk management systems.

ii) Financial software may consist of prepackaged software programs and computer languages tailored to the needs of financial organizations. 
Some funds use open‐source software, and others pay licensing fees for proprietary software.

iii) Financial data providers supply funds primarily with raw financial market data, including security prices, trading information, indices, and company‐specific information such as income statements and balance sheets. 

Due to legal implications, directly marketing alternative investment vehicles can be problematic. 

One method of indirectly marketing private funds is to report a fund's performance to an index provider, especially if the fund's performance is attractive. 

Index providers compile indices of prices that assist fund managers in evaluating performance.

- Consultants

May be hired by allocators to provide a number of roles and services that center on advice, analysis, and investment recommendations.

Offer unbiased analysis of money managers’ investment performance, as well as advice on how to best allocate funds.

Used to serve the role of chief investment officer in small organizations.

The role of an outsourced chief investment officer (OCIO) ranges from performing all of the decision‐making duties of an in‐house chief investment officer to a reduced role of assisting staff with a subset of decisions.

Are compensated for their services in one or more of the following ways: 

i) fees from their clients

ii) fees from proprietary investment products, or 

iii) compensation packages from the money managers for whom they generate business.

Conflicts of interest can emerge when consultants are compensated by money managers because this form of payment can detract from the ability to offer independent advice to clients and encourage the consultant to favor the money managers offering compensation.

Compensation that consultants receive from money managers is undisclosed and can be quite substantial.

Some consultants waive their regular consulting fee, giving the impression that their services are free.

A third compensation approach has emerged in which consultants use their expertise in manager selection and risk management to serve as fund‐of‐funds managers to their clients.

This arrangement avoids explicit fees to the investors for the consulting advice, and offers the potential that the consultants will act with substantial objectivity in the selection of managers.

- Depositories and custodians

Are very similar entities that are responsible for holding their clients’ cash and securities and settling clients’ trades, both of which maintain the integrity of clients’ assets while ensuring that trades are settled quickly.

The Depository Trust Company (DTC) is the principal holding body of securities for traders all over the world and is part of the Depository Trust and Clearing Corporation (DTCC), which provides clearing, settlement, and information services.

- Banks

Commercial bank focuses on the business of accepting deposits and making loans, with modest investment‐related services.

Investment bank focuses on providing sophisticated investment services, including underwriting and raising capital, brokerage services, mergers, and acquisitions.

Hedge funds may enlist the services of a commercial bank to facilitate the flow of both investment‐ and non‐investment‐related capital. 

Hedge funds may also use their commercial bank for loans, credit enhancement, and/or lines of credit.

#### 4) Regulator

#####

Privately organized investment vehicles generally receive reduced regulatory scrutiny because the participants involved tend to be sophisticated institutions or individuals perceived to be less in need of regulatory protection than the general public.

These regulations may be insufficient in protecting the financial markets and/or economy from catastrophic risks, described as systemic (or systematic) risk.

Systemic risk is the potential for economy‐wide losses attributable to failures or concerns over potential failures in financial markets, financial institutions, or major participants.

Given the complexity, opaqueness, use of leverage, and multitude of other factors, the participants in the alternative investments universe have the potential of contributing to systemic risk factors.

\newpage

### Alternative Investment Legal Structures

There are numerous legal structures used throughout investments to meet the preferences and needs of investors and managers.

- Limited Liability and Passive Investments

Limited liability is the restriction of potential loss to a fixed sum.

Receiving this protection in most jurisdictions requires that the investor or manager not be actively engaged in control of the firm's operations.

The issue of liability is therefore related to whether or not the investment is passive.

In the context of investment strategies, passive investments can refer to buy‐and‐hold strategies often designed to match the risk and return of an index.

In the context of limiting liability, passive investments are positions in entities over which the owner of the position does not exert substantial control and therefore may receive reduced liability exposures and/or passive investment tax treatments. 

Limited liability raises concerns over probity in the management of enterprises when investors and managers with limited liability are in the asymmetric position of having huge potential profits and limited losses. 

Probity is the quality of exercising strong principles such as honesty, decency, and integrity. 

- Corporations and Limited Liability Companies

A corporation is a legal entity in which its shareholders receive distributions on a pro rata or pari passu basis and which generally provides limited liability for passive investors.

A limited liability company (LLC) is a distinct entity:

i) Designed to offer its investors protection from losses exceeding their investments absent fraud/other activities that could “pierce the veil” between the member's ownership interest in the LLC and the member's other holdings, and

ii) Does not require distributions and any other advantages of ownership be made in proportion to each member's capital contribution to the firm.

An LLC may be structured to make distributions to members with or without consideration of ownership percentages.

Alternative investments, investors and managers use LLCs and other entities offering limited liability to serve as a buffer to reduce liability exposures.

Entities are created and used for various purposes such as shielding assets, reducing taxation, simplifying regulation, and avoiding entanglements in proceedings such as bankruptcy.

- Limited Partnership Structures

The General Partner (GP) organizes the fund and manages the assets. 

The limited partners contribute most of the capital.

```{r, eval=FALSE}
\begin{center}
\includegraphics[width=10cm, height=3.5cm]{"C:/Catapult Programme/Training/CAIA/Images/2.2.2 Im1.png"}
\end{center}
```

The general partners likely formed an LLC (“parent LLC”) that in turn formed an LLC to serve as the general partner and another LLC to serve as the investment adviser to the partnership.

The personnel involved in three LLCs may be overlapping or even identical, with the primary functions of preserving liability and providing clarity with regard to control.

Each PE fund is managed by a general partner. 

The general partner is typically the PE firm that raised the capital for the fund. 

The general partner sources investment opportunities for the fund, reviews business plans, performs due diligence, and, once an investment is made, takes a seat on the board of directors of the start‐up company and works with the management of the company to develop and implement the business plan.

```{r, eval=FALSE}
\begin{center}
\includegraphics[width=10cm, height=5cm]{"C:/Catapult Programme/Training/CAIA/Images/2.2.2 Im2.png"}
\end{center}
```

At the base of the diagram is the underlying PE investment, which consists of the private business enterprises from which all cash flows to investors must ultimately be derived.

The middle level represents PE funds, which are investment pools created to hold portfolios of PE securities (i.e. equity securities at the bottom).

The funds serve as intermediaries between the underlying business enterprises (called portfolio companies when they are owned by a fund) and the investors in the PE funds.

Institutional investors typically invest in PE as LPs in these funds rather than through direct ownership of PE securities. 

These PE funds are also often referred to as PE investments.

Limited partnership funds (indicated PE funds), are also referred to as PE investments.

- Bankruptcy‐Remote Entities

Is structured as a special purpose vehicle or entity.

A special purpose vehicle (SPV) or special purpose entity (SPE) is a legal entity such as an LLC that serves a specific function (such as holding assets), often with the goal of being bankruptcy remote. 

A bankruptcy‐remote entity is designed to provide protection from involvement in potential bankruptcy proceedings of the entities placing assets in the vehicle.

- Entities Facilitating Investor Taxation Differences

Master‐feeder funds are designed to provide efficient access to investors who are subject to different taxation but wish to invest in the same portfolio.

Offshore jurisdictions typically do not charge taxes on investment income in the country of the fund's domicile, but trust that investors will pay taxes on the investment income in their home country.

Offshore vehicles are designed to be efficient tax‐neutral investments vehicles appropriate for asset owners with a wide array of home countries and tax regimes.

The master fund/master trust is the legal structure (typically offshore) used to invest the assets of both onshore investors and offshore investors in a consistent, if not identical manner, so that both types of investors can share the manager's skills. 

Investors access the master fund through feeder funds.

A feeder fund is a legal structure through which investors with common preferences with regard to taxation can have efficient access to the investment performance of the master trust.

Together, the master trust and feeder funds are referred to as a master‐feeder structure.

```{r, eval=FALSE}
\begin{center}
\includegraphics[width=10cm, height=5cm]{"C:/Catapult Programme/Training/CAIA/Images/2.2.2 Im3.png"}
\end{center}
```

The purpose of the master fund is tax neutrality, not evasion.

For example, in Bermuda, master trust funds pay only a corporate licensing fee, not corporate income tax to ensure that there are no tax consequences to the fund investors at the master trust level. 

\newpage

### Key Features of Fund Structures

- Key Partnership Documents

Key documents used in establishing and managing a hedge fund, private equity fund, or other private partnership: 

i) private‐placement memoranda (or offering memorandum)

Include the formal description of an investment opportunity that complies with securities regulations.

ii) (limited) partnership agreement

A formal written contract creating a partnership.

iii) subscription agreement

An application submitted by an investor who desires to join a limited partnership.

iv) management company operating agreement

An agreement between members related to a limited liability company and the conduct of its business as it pertains to the law.

Limited partnership agreement (LPA) defines its legal framework and its terms and conditions.

Private equity managers are not fiduciaries, so limited partners should not assume that the general partner has the best interest of their investors in mind.

The LPA has two main categories of clauses: 

i) investor protection clauses - cover investment strategy, including possible investment restrictions, key‐person provisions, termination and divorce, the investment committee, the LP advisory committee, exclusivity, and conflicts.

ii) economic terms clauses -  include management fees and expenses, the GP's contribution, and the distribution of cash (i.e. the waterfall)

Distribution waterfall defines how returns are split between the LP and GP and how fees are calculated. 

LPAs are continuously evolving, given the increasing sophistication of fund managers and investors, new regulations, and changing economic environments. 

The LPA lays out conditions aimed at both aligning the interests of fund managers with their investors and discouraging the GP from cheating (moral hazard), lying (adverse selection), or engaging in opportunism (holdup problem) in whatever form. 

Moral hazard and adverse selection take place when there is asymmetric information between two parties.

Adverse selection takes place before a transaction is completed, when the decisions made by one party cause less desirable parties to be attracted to the transaction.

Moral hazard takes place after a transaction is completed and can be defined as the changes in behavior of one or more parties as a result of incentives that come into play once a contract is in effect. 

In economics, the holdup problem is a situation in which two parties refrain from cooperating due to concerns that they might give the other party increased bargaining power and thereby reduce their own profits.

- Corporate Governance in Private Funds

The law and the LPA define and restrict the degree of control LPs have over the activities of GPs.

Such controls relate to waiving or accepting investment restrictions, extending the investment period or fund duration, handling key‐person‐related issues, or participating in an LP advisory committee (LPAC), whose responsibilities are defined in the LPA and normally relate to dealing with conflicts of interest, reviewing valuation methods, and any other consents predefined in the LPA.

LPs can make decisions with either a simple majority or a qualified majority.

A qualified majority is generally more than 75% of LPs in contrast to the over 50% required for a simple majority.

In limited partnership structures, an overactive LP could become reclassified as a GP, thereby losing limited liability.

Another important element of corporate governance is reporting to LPs. 

Various PE associations or industry boards have released guidelines for valuation and reporting.

The obligation to disclose in compliance with these guidelines is increasingly being made part of contractual agreements.

While some GPs reduce the level of detail provided to the bare minimum and share it with all LPs, others share different levels of detail depending on the specific type of investor.

- Investment Objectives, Fund Size, and Fund Terms

In LPAs, the description of investment objectives should be specific but not too narrow.

With uncertain investments, severe information asymmetry, and difficulties in monitoring and enforcing restrictions, fund managers may simply find ways around narrow restrictions.

The fund's size, in terms of capital committed by LPs, needs to be in line with these investment objectives. 

Various other factors implicitly set a minimum or maximum size of the fund.

Fund lives or terms represent a trade‐off among better investment returns, sufficient time to invest and divest, and the degree of illiquidity still acceptable for investors.

The extension of a fund's life is approved annually by a simple majority of LPs or members of the LPAC, one year at a time versus two or more at once, during which time management fees are either reduced or eliminated altogether to stimulate quick exits.

Proceeds are distributed to investors as soon as is feasible after the realization of or distribution of a fund's assets, but LPs grant fund managers the discretion to reinvest some of the proceeds that are realized during the investment period.

- Management Fees and Expenses

In alternative investments, compensation should overwhelmingly be performance driven. 

When LP/GP interests are fully aligned, management fees provide a base compensation so that the fund manager can support the ongoing activities of funds and GPs should not be able to make significant profits on management fees alone. 

- Global Regulations and Fund Structures

The Markets in Financial Instruments Directive (MiFID) is an EU law that establishes uniform regulation for investment managers in the European Economic Area.

The MiFID is one of the primary pieces of European legislation dealing with regulation of investment services, including management services. 

The MiFID II is a revision directed toward extending the reach of MiFID to cover gaps in the 2007 document and addressing emerging issues, such as lack of transparency in trading occurring in dark pools.

In July 2011, the Alternative Investment Fund Managers Directive (AIFMD) came into force, and applies to alternative investment fund managers (AIFMs) that are located in the EU or, if located outside the EU, manage either EU or market funds (whether EU or non‐EU) in the EU.

An AIFM includes any legal or natural person whose regular business is to manage one or more alternative investment funds (AIFs).

An AIF is any collective investment that invests in accordance with a specified policy, except UCITS.

\newpage

### Drawdown Fund Fees and Terms

- Management Fees and Carried Interest

The difference between public and PE governance structures are the compensation structures that provide incentives for the managers and investors

Management fees for PE depend on the size of the fund, generally ranging from 1.5% of committed capital for large funds to 2.5% for smaller funds.

During the investment period, the management fee is assessed on the amount of committed capital, not invested capital. 

Once the investment period concludes, the management fee transitions to being charged on invested capital, not committed capital.

The manager collects a management fee from the moment an investor signs a subscription agreement to invest capital in the fund, even though no capital has actually been contributed by the LP yet. 

The potential for the manager to earn fees on capital from capital calls may give the manager an incentive to call for capital, even when investment opportunities are not of the highest quality.

General partners can earn a wide variety of fees from their portfolio companies, including director fees, buyout fees, financing fees, divestment fees as detailed in the management services agreement between the GP and the portfolio company.

LPs would prefer that all fees earned by the GP would offset the management fees due from the LPs.

The main upside incentive for GPs comes in the form of carried interest, typically 20% of the profits realized by the fund.

Incentive fees are typically entitled carried interest in PE funds and can be calculated on either a fund‐as‐a‐whole or a deal‐by‐deal basis.

- Clawback Provisions

Most PE partnership agreements include a clawback provision.

A clawback provision is a covenant that allows the LPs to receive back (or claw back) previously paid incentive fees. 

The previously paid incentive fees are returned if, at the end or liquidation of the fund, the LPs have returns below the annualized preferred return, or paid an incentive fee on unrealized gains at a level in excess of actual realized gains.

A clawback escrow agreement is where a portion of the manager's incentive fees are held in a segregated account until the entire fund is liquidated. 

This ensures the fund manager does not walk away with incentive fees unless the LPs earn a profit in excess of the hurdle rate or the preferred return.

Only after every limited partner has earned a profit, are the escrow proceeds released to the manager. 

This covenant may stipulate that all management fees must also be recouped by the LPs before the manager can collect incentive fees.

Responsibility for payment of the clawback rests with the persons or entity that received the carried‐interest distributions, then the GP is required to return some proceeds to make the investor whole.

Clawback also exist for LPs, its triggered when at the end of a fund's life, the GPs have received less than their share of the fund's profits.

Such a clawback is relevant where a portion of the committed capital has not been drawn, and LPs have received full repayment of their commitment plus hurdle before the GP has access to their carry.

The solution is that GPs do not take carried interest until all invested, or sometimes even all committed, capital has been repaid to investors.

This could be demotivating as it takes several years for fund's team to see any gains.

Carried interest is usually distributed to GPs as profits are realized.

A clawback ensures that managers will not receive a greater share of the fund's distributions than they are entitled to.

- Carried Interest and Hurdle Rates

Carried interest can be subject to a hurdle rate, or preferred return, so that it begins to accrue only once investors have received their capital back and a minimum previously agreed‐on annual rate of return.

Once preferred return is attained, GPs receive 100% of returns to the point at which they have received the carried interest on the entire amount.

This is called a catch‐up and is synonymous with the soft hurdle concept used by hedge funds.

```{r, eval=FALSE}
\begin{center}
\includegraphics[width=10cm, height=5cm]{"C:/Catapult Programme/Training/CAIA/Images/2.2.2 Im4.png"}
\end{center}
```

In the hurdle zone, the LPs receive all the cash inflows. 

This is followed by the catch‐up period, during which the GP receives all or the major share of distributions. 

These zones can cause unintended consequences, i.e. the catch‐up period ends when the agreed‐on carried‐interest split, or stated share of the fund's profits to be received by the GP and the LPs, is reached. 

Thereafter, distributions are shared between the fund manager and investors in the proportion agreed to in the legal documentation, normally 20% for the GP and 80% for LPs.

Sometimes there is a full catch‐up, whereas other agreements may foresee only a partial catch‐up or no catch‐up at all.

With a 100% catch‐up, the LPs are effectively excluded from all new distributions until the GP has fully recovered the agreed‐on GP share of the profit.

#### Example 1

Assume a €100 million contribution by the LPs in the first year to fund an investment, an 8% hurdle rate, a 100% catch‐up, an 80/20 carry split, and the sale of the investment by the fund in the second year for €200 million. 
The sales proceeds are distributed as below. 

- The €2 million catchup for the GP is found by ensuring that the GP gets 20% of the total, which is 25% of the LPs' payout. 
25% is found as: 20%/(1 – 20%) where 20% is the GP's share.

NB: simply splitting the profit of €100 million (€200 million in sales proceeds minus the €100 million purchase price) on an 80/20 basis would have given the same result.

This is because the fund's return is above the catch‐up zone, and therefore the hurdle is extinguished and does not make any difference.

```{r, eval=FALSE}
\begin{center}
\includegraphics[width=10cm, height=5cm]{"C:/Catapult Programme/Training/CAIA/Images/2.2.2 Im5.png"}
\end{center}
```

#### Example 2

Assume a R200 million contribution by the LPs in the first year to fund an investment, a 6% hurdle rate, a 100% catch‐up, an 80/20 carry split, and the sale of the investment by the fund in the third year for R300 million. 
What is the total cash flow to the LPs and the total cash flow to the GPs? 
Assume the hurdle rate is not compounded and ignore all other expenses such as management fees.

- The preferred return to the LPs is R24 million (R200 million × 6% × 2), leaving R76 million after also returning the R200 million to the LPs. 
The 100% catch‐up to the GPs for an 80%/20% split is R6 million (R24 million × 20%/(1 – 20%)). 
The residual is R70 million, found as R300 million – (R200 million + R24 million + R6 million), and is split 80%/20%.

- Perverse Incentives from Fees

There are two potentially major perverse incentives generated by hurdle rates.

i) High hurdle rates can cause excessive risk taking

Managers of funds with overly high hurdle rates (or of struggling funds) can become unmotivated if it becomes unlikely that they will receive carried interest. 

When the GP's call‐option‐like incentive fee moves far out‐of‐the‐money, the high hurdle rates may lead to desperate risk taking. 

The option‐like payoffs to GPs may mean that the more risk a manager takes, the greater the upside potential, with little immediate downside impact from losses.

Hence, the GP's capital contribution is so important.

ii) There are timing issues with hurdle rates (preferred returns)

GPs face the dilemma of whether to realize an investment over a short period of time to optimize the spread of the IRR over the hurdle rate or to hold on to the investment and try to optimize the long‐term multiple.

The standard preferred return, being based on the IRR, gives incentive to the former.

There has been an increasing use of subscription lines, which are lines of credit used by the GP to make investments in portfolio companies. 

If subscription lines are used to delay capital calls and reduce the time that LPs are invested in the fund, the IRR of the fund may be higher than if capital calls were made at the time the investments were made in the portfolio company.

- GP's Contribution to Initial Private Equity Fund Investment

Excessive risk taking can be reduced or eliminated if managers have a significant portion of their personal wealth in the fund.

Investors in PE funds see 1% of the fund's committed capital as a standard and acceptable contribution for GPs to make. 

The capital contribution by GPs (i.e. hurt money) should be contributed in cash rather than through the waiver of management fees (or as surplus from the management company's budget).

LPs prefer that the GP contribution be at a level that is a substantial portion, thus more than 1% for wealthy managers.

The GP's contribution to the fund is a significant share of their personal wealth, whereas the LPs’ investment, although in absolute terms is far higher, represents an immaterial share of the institution's overall assets.

It is a challenge to determine the appropriate contribution level for a GP: one that provides a reasonable incentive but is not excessively onerous.

- Key‐Person Provision

Depending on the size, experience, and depth of the team, the inability of one or several key persons to carry out their duties could have a substantial adverse effect on the partnership.

The key‐personnel clause is a common and important clause in partnerships that allows LPs to take specified actions in the event that one or several of the named key persons depart the team, stop committing sufficient time to the management of the fund, or sell their interests in the management company.

The key‐personnel (or key‐person) clause (or provision) allows LPs to suspend investment and/or divestment activities until a replacement or replacements are found.

The LPs may even be able to terminate the fund if they so choose.

- Termination and Divorce

LPs may include a bad‐leaver clause, which is a for‐cause removal of the GP that, if exercised (following a simple majority vote of the LPs), causes investments to be suspended until a new fund manager is elected or, in the extreme, the fund is liquidated

In practical terms, conditions leading to a for‐cause removal are very little to legally enforce; issues are highly subjective, and taking matters to court carries high legal risk for an investor, as it's very difficult and lengthy to prove wrongdoing.

A good‐leaver clause enables investors to cease additional funding of the partnership with a vote requiring a qualified majority.

This "without‐cause" clause provides a clear framework for shutting down a partnership that is not working, or when confidence is lost.

To exercise one of these clauses, reputation considerations play a key role, as the market consists of a small number of players who repeatedly interact with one another. 

GPs removed with or without cause may subsequently be unable to raise funds or participate in investment syndicates with other partnerships. 

To avoid such disastrous outcomes, GPs tend to agree on a fund restructuring to prevent a forced removal.

- Other Covenants

The most important covenant is the size of an investment by the fund, usually expressed as a percentage of the capital committed to the fund.

This ensures that the manager does not commit too much capital to a single investment.

Other covenants may include a restriction on the use of debt or leverage by the fund. 

Gearing is a term for increasing risk through leverage. 

In addition, there may be a restriction on co‐investments with prior or future funds controlled by the manager.

There is usually a covenant regarding the distribution of profits, i.e. it's optimal for investors to receive the profits as they are realized.

Distributed profits reduce the amount of committed capital in the fund, which in turn reduces the fees paid to the manager. 

It is in the fund manager's economic interest to retain profits.

Restrictions on the general partner's activities include a limit on the amount of private investments the fund manager can make on its own in any of the firms funded by the fund. 

If the fund manager makes private investments in a select group of companies, these companies may receive more attention than the remaining portfolio of companies funded by the fund. 

GPs are often limited in their ability to sell their general partnership interests in the fund to a third party. 

Such a sale would be likely to reduce the general partner's incentive to monitor investments and produce an effective exit strategy for the fund's portfolio companies.

Another covenant is a restriction on the amount of future fundraising.

Fundraising is time‐consuming and can detract from the manager's time spent managing the investments of the existing fund.

The other covenant is when LPs demand the general partner spend substantially all of the time managing the investments of the fund; outside interests are limited or restricted.

\newpage

### Financial Markets

A primary market refers to the methods, institutions, and mechanisms involved in the placement of new securities to investors. 

A secondary market facilitates trading among investors of previously existing securities.

- Primary Capital Markets

Participants in alternative investments often create securities that are not subsequently listed.

Private equity firms often use primary markets as exit strategies for their underlying investments. 

Large private equity firms hold substantial controlling positions in the companies. 

A goal of these private equity firms is to develop these companies to the point that they can be sold to the public through IPOs.

- Secondary Capital Markets

Market making is a practice whereby an investment bank or another market participant deals securities by regularly offering to buy securities and sell securities.

The market maker seeks to receive the bid‐ask spread through regularly selling at the ask price and buying at the bid price (i.e. bid-ask spread).

Participants that place market orders are market takers.

- Third and Fourth Private Markets

Third markets are regional exchanges where stocks listed in primary secondary markets can also be traded. 

Third markets represent a segment of the OTC market where nonmember investment firms can make markets in and trade securities without going through the exchange.

Fourth markets are electronic exchanges that allow traders to quickly buy and sell exchange‐listed stocks via the electronic communications systems offered by these markets.

Registered broker‐dealers provide sponsorship for these systems so that traders have an alternative system to physical exchanges to buy and sell stocks.

These alternative trading systems are computerized trading systems that do not formally list stocks but include electronic communication networks serving retail brokers and small institutional traders, as well as electronic crossing systems that match large buy and sell orders.

These private financial markets are non‐regulated markets that are neither exchanges nor OTC. 

Much of the high‐frequency trading takes place in the fourth market. 

The advantages of private markets may include:

i) lower transaction costs, 

ii) ease of completing a transaction directly between a buyer and a seller (may/may not involve a broker), and 

iii) ability to expedite the consummation of a transaction

Conversely, the disadvantages may include:

i) existence of asymmetrical information (between the participants)

ii) lack of transparency, and 

iii) lack of regulatory protections

\newpage

### Short Selling

A short position can lead to unlimited losses.

Short positions can generate only limited profits because the stock price cannot fall below zero.

Short selling offers an unattractive profile of limited profit potential and unlimited loss potential, the opposite of long positions.

- The Institutional Mechanics of Short Selling

Financial institutions often lend these securities to short sellers to generate income.

Street name refers to the brokerage practice of having the direct legal ownership of customer securities held in the name of the brokerage firm on behalf of the customers rather than having the legal ownership of the shares reside directly with the economic owners of the securities (i.e., the customers of the brokerage firm).

New shares are not created, but the pool of owners expand through the borrowing and reselling of existing shares.

Substitute dividends are cash flows paid by share borrowers to share lenders to compensate the lenders for the distributions paid by the corporation while the loan of stock is outstanding.

These dividents must be factored into the overall return.

- The Mechanics of Short Selling to the Short Seller

For stocks with a large pool of lenders and high ease of borrowing, the lender will typically pay the broker a portion of the income earned on the collateral. 

This is known as a rebate.

A rebate is a payment of interest to the securities’ borrower on the collateral posted.

A typical rebate for these types of securities is a little less than current short‐term market interest rates.

For stocks in which a small pool of lenders is available and fewer shares are available to borrow the borrower may compensate the lenders to entice them into making the loan.

Here, the cost of borrowing is referred to as a negative rebate or simply the borrow fee rate.

Dividend irrelevancy is the proposition that, in the absence of imperfections such as income taxation that penalized dividends, the distribution of corporate dividends does not alter shareholder wealth.

In practice, this relationship may not hold up and short sellers must recognize they are responsible for the entirety of the dividend irrespective of any change in stock price resulting from the issuance of a dividend by the company.

If stock price rises, the brokerage firm/lender may require the borrower to post additional cash to ensure the shares can be repurchased. 

If the borrower is unable to post cash in a timely manner, the brokerage firm has the right to use existing cash to close out the position and prevent further losses. 

When the short seller is prepared to voluntarily close the position, the short seller uses the posted cash to repurchase shares in the open market.

#### Example 3

Suppose short seller establishes a short position in one share of XYZ at R100 per share and that XYZ pays a dividend of R1.00 per share each year. 
The current rebate on XYZ shares is 1% per year and lenders require loans to be collateralized 102%. 
What would be the rand profit or loss to the short seller if XYZ rose to R103 at the end of one year when the position was liquidated, and what is the effect of the dividend on that profit or loss? 
Assume that commission and other transaction fees per trade totaled 5 cents per share.

- The short position loses R3 (a capital loss). 
When the stock pays dividends of R1, it is up to the short seller to make a cash payment to the securities lender in lieu of dividends, so the short seller loses another R1. 
The institutional short seller received a short stock rebate of R1.02 (1% × R102). 
The commission and other fees total R0.10 across both trades resulting in a total loss of R3.08 per share (–R3.00 + R1.02 – R1.00 – R0.10 commission).

Most securities lending is performed on an overnight basis, wherein securities lenders may demand return of the shares at any time and may require regular adjustment of the collateral amount to reflect the current market price of the borrowed securities. 

Some short sales can be performed as term loans of perhaps six months, wherein the lender agrees not to demand return of the securities until the term has ended. 

- Further Discussion of the Negative Rebate

There are special risks to having short positions in equity securities, especially for stocks that are popular targets of short sellers.

As the quantity of a stock's outstanding shares being lent to short sellers increases, the competition to find new stock to borrow increases.

Entities that hold the stock put that stock "on special."

A special stock is a stock for which higher net fees are demanded when it is borrowed.

To the short seller, this means receiving a smaller or negative rebate, i.e. for general collateral stocks.

When numerous speculators establish highly similar large positions, it is often referred to as a crowded trade.

In the case of traders establishing large short positions, the trade is often termed a crowded short. 

When the inventory of stock available to borrowers becomes extremely tight, short sellers may find their position bought in, meaning the broker revokes the borrowing privilege for that specific stock and requires the trader (i.e., the borrower) to cover the short position.

A short squeeze occurs when holders of short positions are compelled to purchase shares at increasing prices to cover their positions due to limited liquidity. 

As the ratio of shares being sold short increases relative to the total number of freely floating shares, it becomes increasingly difficult to borrow additional shares, and the potential for a short squeeze increases.

Another potentially huge complexity from short selling is that the lender of the security may demand that the shares be returned.

When this happens, the broker simply arranges for another securities lender to loan shares so that the short seller maintains a seamless exposure.

The short seller is forced to close the position during a period of turbulence rather than at a time of their choosing. 


